---
title: "ODI cricket Analysis"
output: html_document
---
setwd("C://Users//Administrator//Desktop//EDAdatasets")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
odi = read.csv("C://Users//Administrator//Desktop//EDAdatasets//odi-batting.csv")
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyverse)
#install.packages("ggmap")
```

library(knitr)


# Cricket Analysis
- [This is hyperlink](http://google.com)
- Missing value analysis
- Univariate Analysis
- Bivariate Analysis


## Data Preview
```{r}
kable(head(odi))
```


### Display the dimensions of data
```{r}
dim(odi)
```


### Display unique countries and also the total number of unique countries
```{r}
a<-(unique(odi$Country))
length(a)
levels(a)
#NROW(a)
#odi%>% select(country)%>% 

```


### Identify how many records are there for Sachin R Tendulkar
```{r}
odi%>%select(Player)%>%filter(Player=='Sachin R Tendulkar')

odi%>% group_by(Player)%>% summarise(total_runs=sum(Runs, na.rm = T)) %>% arrange(-total_runs)%>% head(10)
```


### Top10 players based on total runs
```{r}
head(odi%>%select(Player,Runs)%>%arrange(desc(Runs)),10)
#odi%>%select(Player,Runs)%>%head(arrange(desc(Runs)),10)

```


### Identify top 10 players by total number of matches and filter the data for those ten players and count the number of rows
```{r}
odi%>%select(Player)%>%group_by(Player) %>% summarise(total=n())%>% arrange(desc(total))%>% head(10)

top_players<-odi%>%group_by(Player) %>% summarise(total=n())%>% arrange(-total)%>% head(10)
#odi%>%select(Player,Runs)%>%head(arrange(desc(Runs)),10)
odi%>% filter(Player %in% top_players$Player)%>% nrow()

```


### For top ten players based on maximum runs calculate total runs, average score rate and total number of matches they have played
```{r}
top_players<-odi%>% group_by(Player)%>% summarise(highest=max(Runs,na.rm=T),max_runs=sum(Runs,na.rm =T),avg=mean(ScoreRate,na.rm = T),total=n(),ducks=sum(Runs==0))%>%arrange(-total)%>%head(10)

View(top_players)


odi%>%  group_by(Player) %>% summarise(total=sum(Runs,na.rm=T),avgsr=mean(ScoreRate,na.rm=T), total_matches=n(), ducks=sum(Runs==0, na.rm=T), centuries=sum(Runs>99, na.rm=T))%>% arrange(-total_matches)%>%head(10)
#odi%>% group_by(Player)%>%filter(Player %in% top_players$Player)

```

```{r}
head(odi%>%select(Player,Runs)%>%arrange(desc(Runs)),10)
#odi%>%select(Player,Runs)%>%head(arrange(desc(Runs)),10)

```

```{r}
sum(is.na(odi$Runs)/nrow(odi)*100)
```

```{r}
z<-row.names(odi[is.na(odi$Runs),'Runs'])
```

```{r}
odi[1:10, c('Runs', 'ScoreRate')]
```

```{r}
row.names(odi[is.na(odi$Runs), c('Runs', 'ScoreRate')])
```

```{r}
sapply(odi, is.na)
```

```{r}
is.na(odi)
```


```{r}
sum(is.na(odi$Runs))/nrow(odi)*100
#summarise_each(funs(100*mean(is.na(odi$Runs))))
```
```{r}
#boxplot(odi$Runs)
q1= quantile(odi$Runs, 0.25, na.rm=T)
q2= quantile(odi$Runs, 0.5, na.rm=T)
q3= quantile(odi$Runs, 0.75, na.rm=T)
q4= quantile(odi$Runs, 0.02, na.rm=T)
q5= quantile(odi$Runs, 0.98, na.rm=T)
a<-print(c(q1,q2,q3,q4,q5))
iqr= q5-q4
iqr
lw= q4-1.5*iqr
uw= q5+1.5*iqr
print(c(lw,uw))

outliers=odi%>% filter(Runs<0 | Runs>100)
nrow(outliers) /sum(!is.na(odi$Runs)) * 100
```


### Frequency distribution on categorical column
```{r}
odi %>% group_by(Player)%>%
  summarise(total_matches=n())%>%
  arrange(-total_matches)%>% head(10)%>% ggplot(aes(x=reorder(Player,total_matches),y=total_matches)) + (geom_bar(stat='identity',fill='steelblue')) + (coord_flip())+xlab('Player Name')+ylab('Total No. of Matches')
```


```{r}
q1= quantile(odi$Runs, 0.20, na.rm=T)
q2= quantile(odi$Runs, 0.5, na.rm=T)
q3= quantile(odi$Runs, 0.80, na.rm=T)
a<-print(c(q1,q2,q3))
iqr= q3-q1
iqr
lw= q1-1.5*iqr
uw= q3+1.5*iqr
print(c(lw,uw))

outliers=odi%>% filter(Runs< -52 | Runs>92)
nrow(outliers) /sum(!is.na(odi$Runs)) * 100
```

###Frequency Distribution on categorical column
```{r}
ggplot(odi, aes(x=Country)) + geom_bar(fill='steelblue') + coord_flip()
```


### Percentage of Frequency
```{r}
odi%>% group_by(Player)%>%summarise(totalmatches=n()/nrow(odi)*100)%>%arrange(-totalmatches)%>% ggplot(aes(x=Player,y=totalmatches)) + geom_line(group=1)

```

```{r}
#x=c(1,2,3,4,5,6,7,8,9,10)
#cumsum(x)
odi%>% group_by(Player)%>%summarise(totalmatches=n()/nrow(odi)*100)%>%arrange(-totalmatches)%>% mutate(cum_sum=cumsum(totalmatches)) %>% ggplot(aes(x=reorder(Player,cum_sum), y=cumsum(totalmatches)))+geom_line(group=1) +xlab('player')+ylab('cumsum matches')
```

```{r}
odi$date = as.Date(odi$MatchDate, '%m-%d-%Y')
odi$weekday=weekdays(odi$date)
odi$year=format(odi$date,'%Y')
odi$day=format(odi$date, '%d')
odi$month=format(odi$date, '%m')
odi$weekday=weekdays(odi$date)
odi$quater= quarters(odi$date)

odi%>% filter(Player=='Sachin R Tendulkar')%>% group_by(year)%>% summarise(total=sum(Runs, na.rm = T), avg_runs=mean(Runs, na.rm = T), median_runs=median(Runs, na.rm = T), matches=n())%>% head(25)%>% ggplot(aes(x=year, y=matches))+ geom_line(group=1, color='Red')
```

```{r}
odi%>% filter(Player == 'Virender Sehwag') %>%ggplot( aes(x=Runs, y=ScoreRate)) + geom_point() + geom_smooth()
```

```{r}
odi%>% group_by(Country)%>% summarise(Runs=sum(Runs,na.rm=T), ScoreRate=mean(ScoreRate, na.rm = T))%>%ggplot( aes(x=Runs, y=ScoreRate, color=Country)) + geom_point() + geom_smooth()
```

```{r}
by_player= as.data.frame(sort(table(odi$Player), decreasing = T))
ggplot(head(by_player, 20), aes(x=Var1, y=Freq)) + geom_bar(stat = "identity") + coord_flip()+ labs(x='Player name')
```

```{r}
top_players=odi%>% group_by(Player)%>%summarise(Runs=sum(Runs,na.rm = T))%>%arrange(-Runs)%>% head(5)

odi%>% filter(Player %in% top_players$Player)%>% ggplot(aes(x=Player, y=Runs))+ geom_boxplot()

odi%>% ggplot(aes(x=Country, y=Runs))+ geom_boxplot()
```

