---
title: "EDAassignment3"
author: "Nikhil Shetty"
date: "March 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C://Users//Administrator//Desktop//Datasets//EDAdatasets")
library(knitr)
library(ggplot2)
library(reshape)
Hr_analytics <- read.csv("HR Analytics.csv")

```


#Question 1
Identify percentage of missing values in each column and display the same
```{r}
sapply(Hr_analytics, function(col)sum(is.na(col))/length(col) * 100)
```


#Question 2
Create a function which identify the percentage of outliers in a numerical column and use the function to display the percentage of outliers in all numerical column
```{r}
per_outliers <- function(df){
  column <- names(which(sapply(df, is.numeric)))
  numdf <- df[,column]
  return(sapply(numdf,function(x) length(boxplot.stats(x)$out)/sum(!is.na(x)) * 100)
)}


per_outliers(Hr_analytics)
```


#Question 3
Apply univariate analysis on categorical columns
```{r}
univariate_analysis <- function(df){
  category <- names(which(sapply(df, is.factor)))
  for(i in category){
    print(kable(table(df[,i])))
    print(ggplot(df,aes_string(x=i))+geom_bar())
  }
}

univariate_analysis(Hr_analytics)

```


#Question 4
Apply bivariate analysis like correlation analysis, crosstab analysis and segmented analysis, wherever necessary and share your inferences

#correlation analysis(2 Numerical columns)
```{r}
cor_analysis <- function(df){
  numcol <- names(which(sapply(df, is.numeric)))
  name_comb <- combn(numcol,2,simplify = F)
  result <- c()
  var1 <- c()
  var2 <- c()
  for(col in name_comb){
    if(sd(df[,col[1]]) != 0 & sd(df[,col[2]] != 0)){
      if(sd(df[,col[1]]) > 0 & sd(df[,col[2]] > 0)){
          val <- cor(df[,col[1]],df[,col[2]])
            if(val < 0.05){
                result <- append(result,val)
                var1 <- append(var1,col[1])
                var2 <- append(var2,col[2])
            }
      }
    }
  return(data.frame(x=var1,y=var2,correlation=result))
  }
}
kable(cor_analysis(Hr_analytics))

```


#crosstab analysis(2 categorical columns)
```{r}
crosstab_analysis <- function(df){
  category <- names(which(sapply(df, is.factor)))
  name_comb <- combn(category,2,simplify = F)
  result <- c()
  var1 <- c()
  var2 <- c()
  for(col in name_comb){
    if(nlevels(df[,col[1]]) >=2 & nlevels(df[,col[2]]) >=2){
      val <- chisq.test(df[,col[1]],df[,col[2]])
        if(val$p.value < 0.05){
          result <- append(result,val$p.value)
          var1 <- append(var1,col[1])
          var2 <- append(var2,col[2])
        }
    }
  }
  return(data.frame(variable1=var1,variable2=var2,Pvalue=result))
}
kable(crosstab_analysis(Hr_analytics)) 
```


#Segmented analysis(1 numerical, 1 categorical)
```{r}
segmented <- function(x){
  cat <- names(x)[sapply(x, is.character)]
  num <- names(x)[sapply(x, is.numeric)]
  combntn <- combn(c(num,cat),2,simplify = F)
  a <- c()
  b <- c()
  c <- c()
  d <- c()
  annova <- c()
  t_test <- c()
  i <- 1
  j <- 1
  for (col in combntn){
    if(col[1] %in% cat & col[2] %in% cat){
      next()
    }
    else if(col[1] %in% num & col[2] %in% num){
        next()
    }
    else if(length(unique(x[,col[2]]))>2){
             aov_summary <- aov(x[,col[1]]~x[,col[2]])
             a[i] <- col[1]
             b[i] <- col[2]
             annova[i] <- summary(aov_summary)[[1]][1,"Pr(>F)"]
             i <- i+1
    }
    else if(length(unique(x[,col[2]]))==2 & sd(x[,col[1]])!=0){
            t_test[j] <- t.test(x[,col[1]]~x[,col[2]])$p.value
            c[j] <- col[1]
            d[j] <- col[2]
            j <- j+1
    }
  }
  print(df_aov <- data.frame(a,b,annova))
  print(df_t <- data.frame(c,d,t_test))
}


segmented(Hr_analytics)
```


#Question 5
Include more variables in a charts like scatter plots, multiple box plots, stacked bars etc to identify better insights


Box plot for Department and TotalWorkingYears columns
```{r}
ggplot(Hr_analytics,aes(x=Hr_analytics$Department, y=Hr_analytics$TotalWorkingYears))+ labs(title= "Department vs TotalWorkingYears", x = "Department", y = "TotalWorkingYears")+ geom_boxplot()
```


Scatter plot for age and monthly income columns
```{r}
ggplot(Hr_analytics,aes(x=Hr_analytics$Age, y=Hr_analytics$MonthlyIncome))+ labs(title= "Age vs Monthly income", x = "Age", y = "Monthly income")+ geom_point(stat = "identity")
```


Stacked bar plot for age and number of companies worked columns
```{r}
ggplot(Hr_analytics,aes(x=Age, y=NumCompaniesWorked, fill=Gender))+labs(title= "Age vs NumCompaniesWorked", x = "Age", y = "NumCompaniesWorked")+geom_bar(stat = "identity")
```


Function gives the boxplots for each of the numerical and categorical columns in the dataset.
```{r}
box_plot <- function(df){
  numcol <- names(sapply(df, is.numeric))
  numdf <- df[,numcol]
  catcol <- names(sapply(df, is.factor))
  catdf <- df[,catcol]
print(ggplot(melt(numdf), aes(variable, value)) + geom_boxplot())
print(ggplot(melt(catdf), aes(variable, value)) + geom_boxplot())
}

box_plot(Hr_analytics)
```

